---
title: "Scenario Summaries"
params: 
  scen: scen
  dat_dir: dat_dir
  fit_dir: fit_dir
output: pdf_document
    # rmdformats::html_clean:
    #     code_folding: hide
    #     toc: yes
    #     toc_float: yes
keep_tex: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Packages

```{r}
library(tidyverse)
library(knitr)
library(rstan)
library(devtools)
library(HDInterval)
options(dplyr.summarise.inform = FALSE)
```

## Initializing Functions and Global Objects

```{r}
###################################################
### initialize global objects
###################################################

# manually initialize true effects
blank_eff_mat = matrix(0, nrow = 3, ncol = 4, dimnames = list(c("true_cond_effect", "true_side_effect", "true_int_effect"), c("a1.cond_side", "a2.cond_side", "g.cond_side", "t.cond_side")))

scen1_effect_mat = blank_eff_mat
scen1_effect_mat[1,4] = 0.5

scen2_effect_mat = blank_eff_mat
scen2_effect_mat[2,4] = -0.25

scen3_effect_mat = blank_eff_mat
scen3_effect_mat[1,4] = 0.5
scen3_effect_mat[2,4] = -0.25

scen4_effect_mat = blank_eff_mat
scen4_effect_mat[1,4] = 0.5
scen4_effect_mat[2,4] = -0.25
scen4_effect_mat[3,4] = 0.5

scen5_effect_mat = scen4_effect_mat

scen6_effect_mat = scen4_effect_mat

scen7_effect_mat = blank_eff_mat

true_effect_list = list(scen1_effect_mat,
                        scen2_effect_mat,
                        scen3_effect_mat,
                        scen4_effect_mat,
                        scen5_effect_mat,
                        scen6_effect_mat,
                        scen7_effect_mat)
```


<!-- compiled via slurm sob array submission, ignore below -->
<!-- ```{r} -->
<!-- #set data and fit directories -->
<!-- dat_dir = params$dat_dir -->
<!-- fit_dir = params$fit_dir -->
<!-- ``` -->

<!-- ```{r} -->
<!-- scen = params$scen -->

<!-- #create dataframe with true effect values -->
<!-- effect_true_vals = data.frame(par = colnames(true_effect_list[[scen]]), t(true_effect_list[[scen]])) -->
<!-- rownames(effect_true_vals) = NULL -->
<!-- ``` -->

<!-- ```{r} -->
<!-- print(params) -->

<!-- print(effect_true_vals) -->
<!-- ``` -->

```{r}
scen = 1 #generate report for scenario scen
dat_dir = "../Simulated_Fits/MDND"
fit_dir = "../Simulated_Data/MDND"
```



```{r}
###################################################
### functions for reading in fits and data
###################################################

#function to read in all fitted data for a scenario as a data.frame
read_scen_fits = function(scen, fit_dir, fit_file){
  
  #load fitted model
  samp = readRDS(paste0(fit_dir, "/scenario_", scen, "/", fit_file))
  
  #initialize dataframe for fitted model
  fitted_dat = data.frame()
  
  #loop over parallel samples
  for(i in seq_along(samp$par.samples)){
    
    #check if chain is NA (error in jags sampler, likely infinite density)
    if(any(is.na(samp$par.samples[[i]]))) next
    
    #initialize sub-sample and sub-label vectors
    samp_vec = c()
    label_vec = c()
    num_vec = c()
    
    #loop over parameters
    for(p in seq_along(samp$par.samples[[i]])){
      
      #check if only 1 chains (there should always only be 1 chain)
      if(seq_len(dim(samp$par.samples[[i]][[p]])[3]) != 1) stop(paste0("Parallel sample ", i, ", parameter ", p, ", does not have exactly 1 chain."))
      
      #initialize sub-sample and sub-label vectors
      sub_samp_vec = c()
      sub_label_vec = c()
      sub_num_vec = c()
      
      #loop over parameter vector
      for(v in seq_len(dim(samp$par.samples[[i]][[p]])[1])){
        
        #concatenate samples for this element of the parameter vector
        sub_samp_vec = c(sub_samp_vec, samp$par.samples[[i]][[p]][v,,1])
        
        #concatenate label
        sub_label_vec = c(sub_label_vec, rep(names(samp$par.samples[[i]])[p], dim(samp$par.samples[[i]][[p]])[2]))
        sub_num_vec = c(sub_num_vec, rep(v, dim(samp$par.samples[[i]][[p]])[2]))
      }
      
      samp_vec = c(samp_vec, sub_samp_vec)
      label_vec = c(label_vec, sub_label_vec)
      num_vec = c(num_vec, sub_num_vec)
    }
    fitted_dat = rbind(fitted_dat, data.frame(samp = samp_vec, par = label_vec, par_num = num_vec, chain = i, dataset = strsplit(fit_file, "_")[[1]][1]))
  }
  return(fitted_dat)
}


get_true_vals = function(scen, dat_dir, fit_dir, true_effect_list){
  
  dat_files = list.files(paste0(dat_dir, "/scenario_", scen, '/'))[!grepl("plot", list.files(paste0(dat_dir, "/scenario_", scen, '/')))]
  fit_files = list.files(paste0(fit_dir, "/scenario_", scen, '/'))[!grepl("plot", list.files(paste0(fit_dir, "/scenario_", scen, '/')))]
  dataset_seeds = sapply(strsplit(fit_files, split = c("_")), function(x) x[1])
  
  
  par_dat = data.frame()
  
  true_effect_mat = true_effect_list[[scen]]
  
  for(f in seq_along(dataset_seeds)){
    
    #load fitted model
    dat = readRDS(paste0(dat_dir, "/scenario_", scen, "/", dataset_seeds[f], ".RDS"))
    
    #initialize
    par_vec = c()
    label_vec = c()
    num_vec = c()
    
    #get parameter levels
    par_levels = sapply(strsplit(names(dat$true_pars), "_"), function(x) paste(x[-1], collapse = '_'))
    
    #get parameter types
    par_types = sapply(strsplit(names(dat$true_pars), "_"), function(x) x[1])
    
    #loop over true parameters by level
    for(i in seq_along(par_levels)){
      
      #coerce vectors to matrices to make following code more concise
      if(!is.matrix(dat$true_pars[[i]])) dat$true_pars[[i]] = matrix(dat$true_pars[[i]], 1, 4)
      
      if(par_types[i] == "theta") {
        par_vec = c(par_vec, dat$true_pars[[i]][,1], dat$true_pars[[i]][,2], dat$true_pars[[i]][,3], dat$true_pars[[i]][,4])
        if(par_levels[i] == "cell"){
          label_vec = c(label_vec, 
                        rep(paste0("a1"), nrow(dat$true_pars[[i]])), 
                        rep(paste0("a2"), nrow(dat$true_pars[[i]])),
                        rep(paste0("g"), nrow(dat$true_pars[[i]])),
                        rep(paste0("t"), nrow(dat$true_pars[[i]])))
        } else {
          label_vec = c(label_vec, 
                        rep(paste0("a1.", par_levels[i]), nrow(dat$true_pars[[i]])), 
                        rep(paste0("a2.", par_levels[i]), nrow(dat$true_pars[[i]])),
                        rep(paste0("g.", par_levels[i]), nrow(dat$true_pars[[i]])),
                        rep(paste0("t.", par_levels[i]), nrow(dat$true_pars[[i]])))
        }
        
        num_vec = c(num_vec, rep(seq_len(nrow(dat$true_pars[[i]])), 4))
      } else {
        #transform sd to precision
        # par_vec = c(par_vec, 1 / (dat$true_pars[[i]][,1] ^ 2), 1 / (dat$true_pars[[i]][,2] ^ 2), 1 / (dat$true_pars[[i]][,3] ^ 2), 1 / (dat$true_pars[[i]][,4] ^ 2))
        par_vec = c(par_vec, dat$true_pars[[i]][,1], dat$true_pars[[i]][,2], dat$true_pars[[i]][,3], dat$true_pars[[i]][,4])
        
        
        
        # if(par_levels[i] == "cell"){
        #   label_vec = c(label_vec, 
        #                 rep(paste0("a1.sd"), nrow(dat$true_pars[[i]])), 
        #                 rep(paste0("a2.sd"), nrow(dat$true_pars[[i]])),
        #                 rep(paste0("g.sd"), nrow(dat$true_pars[[i]])),
        #                 rep(paste0("t.sd"), nrow(dat$true_pars[[i]])))
        # } else {
          label_vec = c(label_vec, 
                        rep(paste0("a1.sd.", par_levels[i]), nrow(dat$true_pars[[i]])), 
                        rep(paste0("a2.sd.", par_levels[i]), nrow(dat$true_pars[[i]])),
                        rep(paste0("g.sd.", par_levels[i]), nrow(dat$true_pars[[i]])),
                        rep(paste0("t.sd.", par_levels[i]), nrow(dat$true_pars[[i]])))
        # }
        num_vec = c(num_vec, rep(seq_len(nrow(dat$true_pars[[i]])), 4))
      }
      
    }
    
    #loop through true effects
    effect_names = rownames(true_effect_mat)
    
    for(i in seq_len(dim(true_effect_mat)[1])){
      par_vec = c(par_vec, true_effect_mat[i,1], true_effect_mat[i,2], true_effect_mat[i,3], true_effect_mat[i,4])
      label_vec = c(label_vec, 
                    paste0("a1.", effect_names[i]), 
                    paste0("a2.", effect_names[i]),
                    paste0("g.", effect_names[i]),
                    paste0("t.", effect_names[i]))
      num_vec = c(num_vec, rep(1, 4))
    }
    
    par_dat = rbind(par_dat, data.frame(true_value = par_vec, par = label_vec, par_num = num_vec, dataset = dataset_seeds[f]))
  }
  return(par_dat)
}
```

# Summary Tables

## Load Simulated Data
```{r}
#get dataframe with fitted and true values
scen_true_vals = get_true_vals(scen = scen, dat_dir = dat_dir, fit_dir = fit_dir, true_effect_list = true_effect_list)

#parameters are labeled differently in simulated data.
        #simulated alias = fitted alias:
            #*.sd.pop = *.cond_side
            #*.sd.cond_side = *.animal
            #*.animal = *.cell
scen_true_vals = scen_true_vals %>% 
    group_by(par) %>% 
    mutate(par = ifelse(grepl("\\.sd\\.animal", par), paste0(strsplit(par, split = "\\.")[[1]][1], ".sd.cell"), par)) %>% 
    mutate(par = ifelse(grepl("\\.sd\\.cond_side", par), paste0(strsplit(par, split = "\\.")[[1]][1], ".sd.animal"), par)) %>% 
    mutate(par = ifelse(grepl("\\.sd\\.pop", par), paste0(strsplit(par, split = "\\.")[[1]][1], ".sd.cond_side"), par))
```

## Read in Output From Fitted Models

```{r}
fit_files = list.files(paste0(fit_dir, "/scenario_", scen, '/'))[!grepl("plot", list.files(paste0(fit_dir, "/scenario_", scen, '/')))]
# dataset_seeds = sapply(strsplit(fit_files, split = c("_")), function(x) x[1])

overall_summary_dat = c()
effect_summary_dat = c()
diagnostic_dat = c()

print(length(fit_files))

#loop over simulation runs
for(f in seq_along(fit_files)){
# for(f in 1:2){
  #progress report
  # cat("Loading data for model ", f, " out of ", length(fit_files), "...\n")
  
  #load fitted data
  samp = read_scen_fits(scen, fit_dir, fit_files[f])
  
  #reduce data for overall summary computation
  overall_summary_tmp = samp  %>%
    group_by(par, par_num, dataset) %>%
    summarise(samp_mean = mean(samp),
              LE_Zero = mean(samp < 0),
              GR_Zero = mean(samp > 0),
              LB = hdi(samp)[1],
              UB = hdi(samp)[2]) %>%
    ungroup
  overall_summary_dat = rbind(overall_summary_dat, overall_summary_tmp)
  rm(overall_summary_tmp)
  
  #reduce data for effect computation
  # effect_summary_tmp = samp %>% 
  #   filter(grepl("cond_side", par, par_num)) %>% 
  #   filter(!grepl("tau", par)) %>%
  #   group_by(par, par_num, chain, dataset) %>%
  #   mutate(samp_num = row_number()) %>%
  #   ungroup %>%
  #   pivot_wider(., names_from = par_num, values_from = samp) %>%
  #   select(-samp_num)
  # colnames(effect_summary_tmp)[4:7] = c("00", "10", "01", "11")
  # effect_summary_dat = rbind(effect_summary_dat, effect_summary_tmp)
  # rm(effect_summary_tmp)
  
  #reduce data for diagnostic computation
  par_vec = unique(samp$par)
  
    #loop over parameter
    for(p in seq_along(par_vec)){
        #progress report
        # cat("Getting MCMC diagnostics for parameter ", p, " out of ", length(par_vec), "...\r")

        #initialize parameter storage
        rhat_vec = c()
        ess_vec = c()

        #initialize other stuff
        d = samp %>%
            filter(par == par_vec[p])
        par_num_vec = unique(d$par_num)

        #loop over parameter number (only non-trivial loop if vectorized parameter)
        for(n in par_num_vec){
         d_tmp = d %>%
            filter(par_num == n) %>%
            group_by(chain) %>%
            mutate(samp_id = row_number()) %>%
            ungroup %>%
            pivot_wider(., names_from = chain, values_from = samp) %>%
            select(-par, -par_num, -dataset, -samp_id) %>%
            as.matrix()

        rhat_vec = c(rhat_vec, Rhat(d_tmp))
        ess_vec = c(ess_vec, ess_bulk(d_tmp))
        }
        diagnostic_dat = rbind(diagnostic_dat, data.frame(rhat = rhat_vec, ess = ess_vec, par = par_vec[p], par_num = par_num_vec, dataset = samp$dataset[1]))
    }
  # cat("\n")
  
  rm(samp)
  gc()
}
```


## Get Overall Summary Table
```{r}
# scen_true_vals_collapsed = scen_true_vals %>%
  # mutate(par = ifelse(grepl(".animal", par), sapply(strsplit(par, split = "\\."), function(x) paste(x[-length(x)], collapse = '.')), par))

par_summary_table = overall_summary_dat %>%
  left_join(x = ., y = scen_true_vals, by = c("par", "par_num", "dataset")) %>%
  mutate(bias = samp_mean - true_value,
         rel_bias = (samp_mean - true_value) / true_value,
         coverage = (true_value > LB) & (true_value < UB),
         MSE = bias^2) %>%
  group_by(par, par_num) %>%
  select(-dataset) %>%
  summarise_all(mean)

# par_summary_table
```

## Get Effect Data
```{r}
cutoff = 0.95
effect_dat = overall_summary_dat %>% 
    filter(grepl("_effect", par)) %>%
    filter(!grepl("priorMean", par)) %>%
    filter(!grepl("priorSD", par)) %>%
    mutate(PowFPR = !((0 > LB) & (0 < UB)),
           LE_Zero_cutoff = LE_Zero >= cutoff,
           GR_Zero_cutoff = GR_Zero >= cutoff) %>%
    group_by(par, par_num) %>%
    select(-dataset) %>%
    summarise_all(mean) %>% 
    select(-par_num)
```


## FPR and Power Table for Fitted Models


```{r}
#restructure to make table with FPR/Power for each effect and parameter
pow_fpr_table = effect_dat %>% 
    select(-samp_mean, -LB, -UB, -LE_Zero, -GR_Zero, -LE_Zero_cutoff, -GR_Zero_cutoff) %>% 
    mutate(effect = strsplit(strsplit(par, split = '\\.')[[1]][2], split = "_")[[1]][1],
           par = strsplit(par, split = '\\.')[[1]][1]) %>%
    pivot_wider(., names_from = effect, values_from = PowFPR)

```

```{r}
knitr::kable(pow_fpr_table)
```


## Posterior Effect Probabilities for Fitted Models

```{r}
#restructure to make table with FPR/Power for each effect and parameter
post_prob_table = effect_dat %>% 
    select(-samp_mean, -LB, -UB, -PowFPR) %>% 
    mutate(effect = strsplit(strsplit(par, split = '\\.')[[1]][2], split = "_")[[1]][1],
           par = strsplit(par, split = '\\.')[[1]][1]) %>%
    pivot_wider(., names_from = effect, values_from = c(GR_Zero, LE_Zero)) %>%
    select(LE_Zero_cond, 
           GR_Zero_cond, 
           LE_Zero_side, 
           GR_Zero_side, 
           LE_Zero_int, 
           GR_Zero_int)

```

```{r}
knitr::kable(post_prob_table)
```


## Posterior Effect w/ Cutoff for Fitted Models

```{r}
#restructure to make table with FPR/Power for each effect and parameter
post_prob_cutoff_table = effect_dat %>% 
    select(-samp_mean, -LB, -UB, -PowFPR, -GR_Zero, -LE_Zero) %>% 
    mutate(effect = strsplit(strsplit(par, split = '\\.')[[1]][2], split = "_")[[1]][1],
           par = strsplit(par, split = '\\.')[[1]][1]) %>%
    pivot_wider(., names_from = effect, values_from = c(LE_Zero_cutoff, GR_Zero_cutoff)) %>%
    select(LE_Zero_cutoff_cond, 
           GR_Zero_cutoff_cond, 
           LE_Zero_cutoff_side, 
           GR_Zero_cutoff_side, 
           LE_Zero_cutoff_int, 
           GR_Zero_cutoff_int)

```

```{r}
knitr::kable(post_prob_cutoff_table)
```

<!-- ## Mean and 95% Intervals of True Effects Across All Datasets -->
<!-- ```{r} -->
<!-- effect_tbl = scen_true_vals %>% -->
<!--     filter(!grepl("tau", par) & grepl("cond_side", par)) %>% -->
<!--     pivot_wider(., names_from = par_num, values_from = true_value) %>% -->
<!--     mutate(`00` = `1`, -->
<!--            `10` = `2`, -->
<!--            `01` = `3`, -->
<!--            `11` = `4`) %>% -->
<!--     select(-`1`, -`2`, -`3`, -`4`) %>% -->
<!--     mutate(interaction_effect = `11` - (`01` + `10`) + `00`) %>% -->
<!--     mutate(side_effect = ((`11` + `10` - interaction_effect) / 2) - ((`01` + `00`) / 2)) %>% -->
<!--     mutate(cond_effect = ((`11` + `01` - interaction_effect) / 2) - ((`10` + `00`) / 2)) %>% -->
<!--     group_by(par) %>% -->
<!--     summarise(mean_interaction_effect = mean(interaction_effect), -->
<!--               int_LB = quantile(interaction_effect, c(0.025)), -->
<!--               int_UB = quantile(interaction_effect, c(0.975)), -->
<!--               mean_cond_effect = mean(cond_effect), -->
<!--               cond_LB = quantile(cond_effect, c(0.025)), -->
<!--               cond_UB = quantile(cond_effect, c(0.975)), -->
<!--               mean_side_effect = mean(side_effect), -->
<!--               side_LB = quantile(side_effect, c(0.025)), -->
<!--               side_UB = quantile(side_effect, c(0.975))) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- kable(effect_tbl) -->
<!-- ``` -->

## Effect Summaries for Fitted Models

```{r}
effect_estimate_table = effect_dat %>% 
    select(-PowFPR, -LE_Zero, -GR_Zero, -LE_Zero_cutoff, -GR_Zero_cutoff) %>% 
    group_by(par) %>%
    mutate(effect = strsplit(strsplit(par, split = '\\.')[[1]][2], split = "_")[[1]][1],
           par = strsplit(par, split = '\\.')[[1]][1]) %>%
    ungroup() %>%
     pivot_wider(., names_from = effect, values_from = c(samp_mean, LB, UB)) %>%
  left_join(x = ., y = (effect_true_vals %>% group_by(par) %>% mutate(par = strsplit(par, split = '\\.')[[1]][1])), by = "par") %>%
    mutate(cond_bias = samp_mean_cond - true_cond_effect,
           side_bias = samp_mean_side - true_side_effect,
           int_bias = samp_mean_int - true_int_effect)


```

```{r}
knitr::kable(effect_estimate_table)
```

## Cond/Side Level Parameter Summaries for Fitted Model

```{r}
par_summary_table %>% filter(!grepl("tau", par) & grepl("cond_side", par)) %>% kable

### CI for Cond/Side Parameters across all simulated datasets
```

## Cond/Side Level Parameter Summaries for Simulated Data

```{r}
scen_true_vals %>%
    filter(!grepl("tau", par) & grepl("cond_side", par)) %>%
    group_by(par, par_num) %>%
    summarise(sim_mean = mean(true_value), LB = quantile(true_value, probs = 0.05), UB = quantile(true_value, probs = 0.975)) %>%
    kable
```


## Precision Tables

### a1

```{r}
## a1
par_summary_table %>% filter(grepl("a1.sd", par)) %>% kable
```

### a2

```{r}
## a2
par_summary_table %>% filter(grepl("a2.sd", par)) %>% kable
```

### g

```{r}
## g
par_summary_table %>% filter(grepl("g.sd", par)) %>% kable
```

### t

```{r}
## t
par_summary_table %>% filter(grepl("t.sd", par)) %>% kable
```

## General Summary of All Parameters

```{r}
knitr::kable(par_summary_table)
```

# Sampler Diagnostics

## Rhat Summaries Across Datasets
```{r}
#seperate cell level parameters, animal level parameters and other parameters from eachother
diagnostic_dat_cell = diagnostic_dat %>% filter(!grepl("\\.", par))
diagnostic_dat_else = diagnostic_dat %>% filter(grepl("\\.", par))
diagnostic_dat_animal = diagnostic_dat_else %>% filter(grepl("animal", par))
diagnostic_dat_else = diagnostic_dat_else %>% filter(!grepl("animal", par))
```

### Table for Cell-level parameters
```{r}
diagnostic_dat_cell %>% group_by(par) %>% 
    summarise(Min = min(rhat), 
              `1st Qu.` = quantile(rhat, probs = 0.25), 
              Mean = mean(rhat), 
              Median = median(rhat), 
              `3rd Qu.` = quantile(rhat, probs = 0.75), 
              Max = max(rhat)) %>%
    kable
```

### Table for Animal-level parameters
```{r}
diagnostic_dat_animal %>% group_by(par) %>% 
    summarise(Min = min(rhat), 
              `1st Qu.` = quantile(rhat, probs = 0.25), 
              Mean = mean(rhat), 
              Median = median(rhat), 
              `3rd Qu.` = quantile(rhat, probs = 0.75), 
              Max = max(rhat)) %>%
    kable
```

### Table for Other Parameters
```{r}
diagnostic_dat_else %>% group_by(par, par_num) %>% 
    summarise(Min = min(rhat), 
              `1st Qu.` = quantile(rhat, probs = 0.25), 
              Mean = mean(rhat), 
              Median = median(rhat), 
              `3rd Qu.` = quantile(rhat, probs = 0.75), 
              Max = max(rhat)) %>%
    kable
```

## ESS Summaries Across Datasets

### Table for Cell-level parameters
```{r}
diagnostic_dat_cell %>% group_by(par) %>% 
    summarise(Min = min(ess), 
              `1st Qu.` = quantile(ess, probs = 0.25), 
              Mean = mean(ess), 
              Median = median(ess), 
              `3rd Qu.` = quantile(ess, probs = 0.75), 
              Max = max(ess)) %>%
    kable
```

### Table for Animal-level parameters
```{r}
diagnostic_dat_animal %>% group_by(par) %>% 
    summarise(Min = min(ess), 
              `1st Qu.` = quantile(ess, probs = 0.25), 
              Mean = mean(ess), 
              Median = median(ess), 
              `3rd Qu.` = quantile(ess, probs = 0.75), 
              Max = max(ess)) %>%
    kable
```

### Table for Other Parameters
```{r}
diagnostic_dat_else %>% group_by(par, par_num) %>% 
    summarise(Min = min(ess), 
              `1st Qu.` = quantile(ess, probs = 0.25), 
              Mean = mean(ess), 
              Median = median(ess), 
              `3rd Qu.` = quantile(ess, probs = 0.75), 
              Max = max(ess)) %>%
    kable
```



